---
title: "Dirichlet"
output:
  html_document:
    code_folding: hide
    collapsed: no
    highlight: haddock
    keep_md: no
    theme: cerulean
    toc: yes
    toc_float:
      collapsed: yes
    toc_depth: 6
    number_sections: false
params:
    date: !r format(Sys.Date(), "%d %B, %Y")
    client: Goldfinch Bio

---

<!-- Indent the TOC -->
<script>
$(document).ready(function() {
  $items = $('div#TOC li');
  $items.each(function(idx) {
    num_ul = $(this).parentsUntil('#TOC').length;
    $(this).css({'text-indent': num_ul * 10, 'padding-left': 0});
  });

});
</script>

<table width=100%>
<tr>
<td>
<h3>Prepared for `r params$client` by:</h3>
Diamond Age Data Science
</td>
<td width=150></td>
<td width=150 align=right>`r params$date`</td>
</tr>
</table>




```{r setup, include=FALSE}
DOCNAME = "in_vivo_dirichlet"
knitr::opts_chunk$set(autodep        = TRUE,
                      cache          = FALSE,
                      cache.path     = paste0("cache/", DOCNAME, "/"),
                      fig.path       = paste0("cache/", DOCNAME, "/"),
                      cache.comments = FALSE,
                      cache.lazy     = FALSE,
                      echo           = TRUE,
                      error          = FALSE,
                      fig.align      = "center",
                      fig.width      = 10,
                      fig.height     = 6,
                      message        = FALSE,
                      warning        = FALSE)

#Libraries
library(here)
library(viridis)
library(knitr)
library(DT)
library(kableExtra)
library(parallel)
library(DirichletReg)
library(Seurat)
library(data.table)
library(RColorBrewer)
library(tidyverse)

source(here("R/utilities.R"))
source(here("R/plot.r"))
source(here("R/colors.r"))

rename = dplyr::rename


#Options
save = TRUE



# Directories
baseDir = here()
saveDir = file.path(baseDir,"output", DOCNAME)
if(!dir.exists(saveDir)){
  dir.create(saveDir)
}



```



```{r}
### from smillie et al
### https://github.com/cssmillie/ulcerative_colitis

dirichlet_regression = function(counts, covariates, formula){

    # Dirichlet multinomial regression to detect changes in cell frequencies
    # formula is not quoted, example: counts ~ condition
    # counts is a [samples x cell types] matrix
    # covariates holds additional data to use in the regression
    #
    # Example:
    # counts = do.call(cbind, tapply(seur@data.info$orig.ident, seur@ident, table))
    # covariates = data.frame(condition=gsub('[12].*', '', rownames(counts)))
    # res = dirichlet_regression(counts, covariates, counts ~ condition)
    
    # Calculate regression
    counts = as.data.frame(counts)
    counts$counts = DR_data(counts)
    data = cbind(counts, covariates)
    fit = DirichReg(counts ~ condition, data)
    
    # Get p-values
    u = summary(fit)
    pvals = u$coef.mat[grep('Intercept', rownames(u$coef.mat), invert=T), 4]
    v = names(pvals)
    pvals = matrix(pvals, ncol=length(u$varnames))
    rownames(pvals) = gsub('condition', '', v[1:nrow(pvals)])
    colnames(pvals) = u$varnames
    fit$pvals = pvals
    
    fit
}
```



```{r, fig.height=4, fig.width=6}

if(file.exists(file.path(saveDir, "seur.RDS"))){
  seur <- read_rds(file.path(saveDir, "seur.RDS"))
}else{
  
  conditions <- c(  "Day_10_2wk_transplant",
                  "Day_10_4wk_transplant",
                  "Day_12_2wk_transplant",
                  "Day_12_4wk_transplant",
                  "Day_14_2wk_transplant",
                  "Day_14_4wk_transplant")
  names(conditions) <- conditions
  
  
  load_seurats <- function(c){
    temp_seur <- read_rds(here("output", c, paste0(c, ".RDS")))
    Idents(temp_seur) <- "orig.ident"
    return(temp_seur)
  }
  
  seur_list <- mclapply(conditions, load_seurats, mc.cores = 6)


  seur <- merge(seur_list[[1]], c(seur_list[[2]], seur_list[[3]], seur_list[[4]], seur_list[[5]], seur_list[[6]]))
  
  rm(seur_list)
  gc()
  
  saveDat(seur)
}

  # seur$condition <- str_split(seur$orig.ident, "_transplant", n=2, simplify = T)[,1]
seur$cell_type_simple <- factor(seur$cell_type_simple, levels = cell_levels_collapse)
  

table(seur$cell_type_simple)

```

# Condensed Cell Types

## Regression by week

```{r, results="hide"}
# counts = do.call(cbind, tapply(seur$orig.ident, seur$cell_type_simple_2, table))
counts = table(seur$orig.ident, seur$cell_type_simple)
counts <- counts[,colSums(counts) != 0]
counts <- ifelse(counts ==0, 1, counts)
covariates = data.frame(condition=str_split(rownames(counts), pattern = "_", n=4, simplify = T)[,3])
  
pvals = dirichlet_regression(counts=counts, covariates=covariates, formula=counts ~ condition)$pvals  
colnames(pvals) <- colnames(counts)
pct = 100*counts/rowSums(counts)
prop = counts/rowSums(counts)

p1 = matrix_barplot(prop, group_by=covariates$condition, pvals=pvals, colors = c("lightgrey", "grey40"), ylab = "Proportion", legend.title = "")


saveDat(prop, filename = "regress_by_week", csv=T)
```

```{r fig.height=2.2, fig.width=3.75}
p1 <- p1 + theme_default() %+replace% theme(axis.text.x = element_text(angle = 90, hjust = 1),
                                            legend.text = element_text(size=6),
                                            legend.key.size = unit(6, "pt"),
                                             axis.text = element_text(size=6),
                                             strip.text = element_text(size = 6, margin = margin(0.1,0,0.1,0, "cm")),
                                             axis.title = element_text(size = 6))
p1
# cowplot::save_plot(p1,filename = "journal_submission/fig2c.pdf", base_height = 2.11, base_width = 3.52)
```

### p values
```{r}
downloadableDT(pvals, rownames = T)

```

### Proportions

```{r}
rns <- rownames(prop)

df_prop <- as_tibble(prop) %>% mutate(sample = rns) %>% separate(sample, into = c("condition", "rep"), sep = "_rep")

df_prop <- df_prop %>% pivot_longer(names_to = "Cell Type", cols = c(colnames(prop))) %>% group_by(condition, `Cell Type`) %>% summarise(mean = mean(value), sd = sd(value))


downloadableDT(df_prop, rownames = T)
```


## Regression by day


```{r, results="hide"}
# counts = do.call(cbind, tapply(seur$orig.ident, seur$cell_type_simple_2, table))
counts = table(seur$orig.ident, seur$cell_type_simple_2)
counts <- counts[,colSums(counts) != 0]
counts <- ifelse(counts ==0, 1, counts)
covariates = data.frame(condition= paste0(str_split(rownames(counts), pattern = "_", n=4, simplify = T)[,1], str_split(rownames(counts), pattern = "_", n=4, simplify = T)[,2]))
  
pvals = dirichlet_regression(counts=counts, covariates=covariates, formula=counts ~ condition)$pvals  
colnames(pvals) <- colnames(counts)
pct = 100*counts/rowSums(counts)

p2 = matrix_barplot(pct, group_by=covariates$condition, pvals=pvals, colors = set.colors, ylab = "Proportion", legend.title = "")


downloadableDT(pvals, rownames = T)
```

```{r}
p2 + theme_default() %+replace% theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust=1.5))
cowplot::save_plot(p2,filename = paste0(saveDir, "/p2.pdf"))
```

```{r}
downloadableDT(pvals, rownames = T)

```

## Regression by day for 2wk

```{r, results="hide"}
# counts = do.call(cbind, tapply(seur$orig.ident, seur$cell_type_simple_2, table))
counts = table(seur$orig.ident, seur$cell_type_simple_2)
counts <- counts[,colSums(counts) != 0]
counts <- ifelse(counts ==0, 1, counts)
counts <- counts[grepl("2wk", rownames(counts)),]
covariates = data.frame(condition= paste0(str_split(rownames(counts), pattern = "_", n=4, simplify = T)[,1], str_split(rownames(counts), pattern = "_", n=4, simplify = T)[,2]))
  
pvals = dirichlet_regression(counts=counts, covariates=covariates, formula=counts ~ condition)$pvals  
colnames(pvals) <- colnames(counts)
pct = 100*counts/rowSums(counts)

p3 = matrix_barplot(pct, group_by=covariates$condition, pvals=pvals, colors = set.colors)


downloadableDT(pvals, rownames = T)
```


```{r}
p3 + theme_bw(base_size = 14) %+replace% theme(axis.text.x = element_text(angle = -45, hjust = 0, vjust=1.5))
cowplot::save_plot(p3,filename = paste0(saveDir, "/p3.pdf"))
```

```{r}
downloadableDT(pvals, rownames = T)

```

## Regression by day for 4wk

```{r, results="hide"}
# counts = do.call(cbind, tapply(seur$orig.ident, seur$cell_type_simple_2, table))
counts = table(seur$orig.ident, seur$cell_type_simple_2)
counts <- counts[,colSums(counts) != 0]
counts <- ifelse(counts == 0, 1, counts)
counts <- counts[grepl("4wk", rownames(counts)),]
covariates = data.frame(condition= paste0(str_split(rownames(counts), pattern = "_", n=4, simplify = T)[,1], str_split(rownames(counts), pattern = "_", n=4, simplify = T)[,2]))
  
pvals = dirichlet_regression(counts=counts, covariates=covariates, formula=counts ~ condition)$pvals  
colnames(pvals) <- colnames(counts)
pct = 100*counts/rowSums(counts)

p4 = matrix_barplot(pct, group_by=covariates$condition, pvals=pvals, colors = set.colors)


downloadableDT(pvals, rownames = T)
```


```{r}
p4 + theme_bw(base_size = 14) %+replace% theme(axis.text.x = element_text(angle = -45, hjust = 0, vjust=1.5))
cowplot::save_plot(p4,filename = paste0(saveDir, "/p4.pdf"))
```

```{r}
downloadableDT(pvals, rownames = T)

```


# With Off Targets Grouped



```{r}
off_targets <- c("Glial", "Melanocyte", "Cartilage", "Muscle")

seur$cell_type_off_targets <- ifelse(as.character(seur$cell_type_simple) %in% off_targets, "Off Target", as.character(seur$cell_type_simple))

cell_levels_off_target <- c(cell_levels_collapse[-which(cell_levels_collapse %in% off_targets)], "Off Target")

seur$cell_type_off_targets <- factor(seur$cell_type_off_targets, levels = cell_levels_off_target)



```



## Regression by week

```{r, results="hide"}
# counts = do.call(cbind, tapply(seur$orig.ident, seur$cell_type_simple_2, table))
counts = table(seur$orig.ident, seur$cell_type_off_targets)
counts <- counts[,colSums(counts) != 0]
counts <- ifelse(counts ==0, 1, counts)
covariates = data.frame(condition=str_split(rownames(counts), pattern = "_", n=4, simplify = T)[,3])


pvals = dirichlet_regression(counts=counts, covariates=covariates, formula=counts ~ condition)$pvals  
colnames(pvals) <- colnames(counts)
pct = 100*counts/rowSums(counts)
prop = counts/rowSums(counts)

p5 = matrix_barplot(prop, group_by=covariates$condition, pvals=pvals, colors = c("lightgrey", "grey40"), ylab = "Proportion", legend.title = "", asterix.size = 3)

downloadableDT(pvals, rownames = T)

saveDat(prop, filename = "collapse_off_target_regress_by_week", csv=T)
```


```{r fig.height=2.5, fig.width=4.25}

p5 <- p5 + theme_default() %+replace% theme(axis.text.x = element_text(angle = 90, hjust = 1),
                                            legend.text = element_text(size=6),
                                            legend.key.size = unit(6, "pt"),
                                            legend.margin=margin(0,0,0,0),
                                            legend.justification = "left",
                                            legend.spacing.x = unit(.1, "char"),
                                            legend.box.margin=margin(-5,-5,-5,-12),
                                            axis.text = element_text(size=6),
                                            strip.text = element_text(size = 6, margin = margin(0.1,0,0.1,0, "cm")),
                                            axis.title = element_text(size = 6))
p5
cowplot::save_plot(p5,filename = "journal_submission/fig2c.pdf", base_height = 2.5, base_width = 4.25)
ggsave("journal_submission/fig2c.png", device="png", width = 4.25, height = 2.5, dpi=600)
```

### p values
```{r}
downloadableDT(pvals, rownames = T)

```

### Proportions

```{r}
rns <- rownames(prop)

df_prop <- as_tibble(prop) %>% mutate(sample = rns) %>% separate(sample, into = c("condition", "rep"), sep = "_rep")

df_prop <- df_prop %>% pivot_longer(names_to = "Cell Type", cols = c(colnames(prop))) %>% group_by(condition, `Cell Type`) %>% summarise(mean = mean(value), sd = sd(value))


downloadableDT(df_prop, rownames = T)
```


## Regression by day


```{r, results="hide"}
# counts = do.call(cbind, tapply(seur$orig.ident, seur$cell_type_simple_2, table))
counts = table(seur$orig.ident, seur$cell_type_off_targets)
counts <- counts[,colSums(counts) != 0]
counts <- ifelse(counts ==0, 1, counts)
covariates = data.frame(condition= paste0(str_split(rownames(counts), pattern = "_", n=4, simplify = T)[,1], str_split(rownames(counts), pattern = "_", n=4, simplify = T)[,2]))
  
pvals = dirichlet_regression(counts=counts, covariates=covariates, formula=counts ~ condition)$pvals  
colnames(pvals) <- colnames(counts)
pct = 100*counts/rowSums(counts)

p6 = matrix_barplot(pct, group_by=covariates$condition, pvals=pvals, colors = set.colors)

```

```{r}
p6 + theme_bw(base_size = 14) %+replace% theme(axis.text.x = element_text(angle = -45, hjust = 0, vjust=1.5))
cowplot::save_plot(p2,filename = paste0(saveDir, "/p2.pdf"))

downloadableDT(pvals, rownames = T)
```

```{r}
downloadableDT(pvals, rownames = T)

```

## Regression by day for 2wk

```{r, results="hide"}
# counts = do.call(cbind, tapply(seur$orig.ident, seur$cell_type_simple_2, table))
counts = table(seur$orig.ident, seur$cell_type_off_targets)
counts <- counts[,colSums(counts) != 0]
counts <- ifelse(counts ==0, 1, counts)
counts <- counts[grepl("2wk", rownames(counts)),]
covariates = data.frame(condition= paste0(str_split(rownames(counts), pattern = "_", n=4, simplify = T)[,1], str_split(rownames(counts), pattern = "_", n=4, simplify = T)[,2]))
  
pvals = dirichlet_regression(counts=counts, covariates=covariates, formula=counts ~ condition)$pvals  
colnames(pvals) <- colnames(counts)
pct = 100*counts/rowSums(counts)

p7 = matrix_barplot(pct, group_by=covariates$condition, pvals=pvals, colors = set.colors)

downloadableDT(pvals, rownames = T)
```


```{r}
p7 + theme_bw(base_size = 14) %+replace% theme(axis.text.x = element_text(angle = -45, hjust = 0, vjust=1.5))
cowplot::save_plot(p3,filename = paste0(saveDir, "/p3.pdf"))
```

```{r}
downloadableDT(pvals, rownames = T)

```

## Regression by day for 4wk

```{r, results="hide"}
# counts = do.call(cbind, tapply(seur$orig.ident, seur$cell_type_simple_2, table))
counts = table(seur$orig.ident, seur$cell_type_off_targets)
counts <- counts[,colSums(counts) != 0]
counts <- ifelse(counts == 0, 1, counts)
counts <- counts[grepl("4wk", rownames(counts)),]
covariates = data.frame(condition= paste0(str_split(rownames(counts), pattern = "_", n=4, simplify = T)[,1], str_split(rownames(counts), pattern = "_", n=4, simplify = T)[,2]))
  
pvals = dirichlet_regression(counts=counts, covariates=covariates, formula=counts ~ condition)$pvals  
colnames(pvals) <- colnames(counts)
pct = 100*counts/rowSums(counts)

p8 = matrix_barplot(pct, group_by=covariates$condition, pvals=pvals, colors = set.colors)

downloadableDT(pvals, rownames = T)
```


```{r}
p8 + theme_bw(base_size = 14) %+replace% theme(axis.text.x = element_text(angle = -45, hjust = 0, vjust=1.5))
cowplot::save_plot(p4,filename = paste0(saveDir, "/p4.pdf"))
```


```{r}
downloadableDT(pvals, rownames = T)

```