---
title: "Pseudotime analysis - in vivo podocytes"
output:
  html_document:
    code_folding: hide
    collapsed: no
    highlight: haddock
    keep_md: no
    theme: cerulean
    toc: yes
    toc_float:
      collapsed: yes
    toc_depth: 6
    number_sections: false
params:
    date: !r format(Sys.Date(), "%d %B, %Y")
    client: Goldfinch Bio

---

<!-- Indent the TOC -->
<script>
$(document).ready(function() {
  $items = $('div#TOC li');
  $items.each(function(idx) {
    num_ul = $(this).parentsUntil('#TOC').length;
    $(this).css({'text-indent': num_ul * 10, 'padding-left': 0});
  });

});
</script>

<table width=100%>
<tr>
<td>
<h3>Prepared for `r params$client` by:</h3>
Diamond Age Data Science
</td>
<td width=150></td>
<td width=150 align=right>`r params$date`</td>
</tr>
</table>



```{r setup, include=FALSE}
DOCNAME = "in_vivo_podocytes2"
knitr::opts_chunk$set(autodep        = TRUE,
                      cache          = FALSE,
                      cache.path     = paste0("cache/", DOCNAME, "/"),
                      fig.path       = paste0("cache/", DOCNAME, "/"),
                      cache.comments = FALSE,
                      cache.lazy     = FALSE,
                      echo           = TRUE,
                      error          = FALSE,
                      fig.align      = "center",
                      fig.width      = 10,
                      fig.height     = 8,
                      message        = FALSE,
                      warning        = FALSE)

#Libraries
library(Seurat)
library(parallel)
library(future)
library(here)
library(GSVA)
library(clustree)
library(viridis)
library(knitr)
library(DT)
library(kableExtra)
library(tidyverse)
library(scales)
library(harmony)
library(monocle)
library(slingshot)
library(multcompView)
library(broom)

source(here("R/utilities.R"))

#Options
nCores <- 12
options(future.globals.maxSize = 8 * 1024^3)
save = TRUE



# Functions



# Directories
baseDir = here()
saveDir = file.path(baseDir,"output", DOCNAME)
if(!dir.exists(saveDir)){
  dir.create(saveDir)
}



```


# Trajectory analysis of Podocytes and Nephron progenitors

In this analysis I combined the podocytes and nephron progenitors from all of the in vivo datasets. I also included podocytes from adult and fetal kidney (same data included in the podocyte correlations).



```{r}

if(file.exists(file.path(saveDir, "seur.RDS"))){
  seur <- read_rds(file.path(saveDir, "seur.RDS"))
}else{
  
  conditions <- c(  "Day_10_2wk_transplant",
                  "Day_10_4wk_transplant",
                  "Day_12_2wk_transplant",
                  "Day_12_4wk_transplant",
                  "Day_14_2wk_transplant",
                  "Day_14_4wk_transplant", 
                  "human_qc")
  names(conditions) <- conditions
  
  
  select_podos <- function(c){
    temp_seur <- read_rds(here("output", c, paste0(c, ".RDS")))
    temp_seur$cell_type_manual_podo <- ifelse(temp_seur$cell_type_manual == "Podocyte Precursor", "Podocyte", temp_seur$cell_type_manual)
    temp_seur$cell_rep <- paste0(temp_seur$orig.ident, "_", temp_seur$cell_type_manual_podo)
    podos <- subset(temp_seur, cell_type_manual %in% c("Podocyte", "Nephron Progenitor"))
    Idents(podos) <- "orig.ident"
    return(podos)
  }
  
  seur_list <- mclapply(conditions, select_podos, mc.cores = 8)


  seur <- merge(seur_list[[1]], c(seur_list[[2]], seur_list[[3]], seur_list[[4]], seur_list[[5]], seur_list[[6]], seur_list[[7]]))
  seur <- subset(seur, orig.ident != "BTN-PH" & orig.ident != "BTN-YY")
  
  rm(seur_list)
  gc()
    
    
  plan(strategy="multiprocess", workers = 8)
  seur <- NormalizeData(seur)
  seur <- ScaleData(seur, vars.to.regress = c("nCount_RNA", "nFeature_RNA"))
  seur <- FindVariableFeatures(seur, selection.method = "vst", nfeatures = 2000)
  seur <- RunPCA(seur, npcs=100)
  # ElbowPlot(seur, ndims = 100)
  
  seur$condition <- ifelse(grepl("_rep", seur$orig.ident), str_split(seur$orig.ident, "_rep", simplify = T)[,1], seur$orig.ident)
  seur$condition2 <- ifelse(grepl("Adult",seur$condition), "AdultKidney", 
                                  ifelse(grepl("Fetal", seur$condition), "FetalKidney", seur$condition))

  seur <- RunUMAP(seur, reduction = "pca", dims = 1:100)
  
  plan(strategy = "sequential")
  
  seur <- FindNeighbors(seur, dims = 1:100)
  seur <- FindClusters(seur, resolution = c(0.2,0.3))
  
  
  
  condition_levels <- c("Adult Kidney", "Fetal Kidney", "D10-2wk", "D10-4wk", "D12-2wk", "D12-4wk", "D14-2wk", "D14-4wk")
  
  conditionDF <- seur@meta.data %>% mutate(condition2 = gsub("day_", "D", condition2),
                            condition2 = gsub("_2wk_transplant", "-2wk", condition2),
                            condition2 = gsub("_4wk_transplant", "-4wk", condition2),
                            condition2 = gsub("Kidney", " Kidney", condition2),
                            condition2 = factor(condition2, levels = condition_levels))
  seur$condition2 <- conditionDF$condition2

  saveDat(seur)
}



```



### Trajectory 2 {.tabset}

Genes that vary most between clusters


```{r, fig.height=6}
if(file.exists(file.path(saveDir, "podo_cds2.RDS"))){
  podo_cds2 <- read_rds(file.path(saveDir, "podo_cds2.RDS"))
}else{
  
  
  clustering_DEG_genes <- differentialGeneTest(podo_cds[expressed_genes,],
                                               fullModelFormulaStr = "~RNA_snn_res.0.3 + batch",
                                               reducedModelFormulaStr = "~batch",
                                               cores = 16)
  
  ordering_genes2 <-
      row.names(clustering_DEG_genes)[order(clustering_DEG_genes$qval)][1:1000]
  
  podo_cds2 <-
      setOrderingFilter(podo_cds,
          ordering_genes = ordering_genes2)
  
  podo_cds2 <-
      reduceDimension(podo_cds2, method = 'DDRTree', verbose = T)
  
  podo_cds2 <-
      orderCells(podo_cds2, reverse = T)
  
  # podo_cds2 <- orderCells(podo_cds2, root_state = 3)
  
  # podo_cds2$condition2 <- conditionDF$condition2
  
  saveDat(podo_cds2)
}


podo_cds2$cell_type_manual_factor <- factor(podo_cds2$cell_type_manual, levels = cell_levels_collapse)
table(podo_cds2$cell_type_manual_factor)
```




#### Labeled by cell type

```{r}
p <- plot_cell_trajectory(podo_cds2, color_by = "cell_type_manual_factor", cell_size = 0.5) +
  scale_fill_manual(values = colors, drop = FALSE, name="") +
  labs(color="", shape="") +
  theme_classic(base_size = 16, base_family = "Roboto") %+replace% theme(legend.position="top",
                                                                         legend.text = element_text(size=16),
                                                                         axis.title = element_text(size=16),
                                                                         axis.text = element_text(size = 16))

p + guides(color = guide_legend(override.aes = list(size = 2)))
```

#### Labeled by cell state

```{r}
plot_cell_trajectory(podo_cds2, color_by = "State", cell_size = 0.5)
```

#### Labeled by Pseudotime

```{r}

plot_cell_trajectory(podo_cds2, color_by = "Pseudotime", cell_size = 0.5) +
  scale_color_viridis_c(name="") +
  guides(color = guide_colourbar(barwidth = 15)) +
  theme_classic(base_size = 16, base_family = "Roboto") %+replace% theme(legend.position="top",
                                                                        legend.text = element_text(size=14),
                                                                        axis.title = element_text(size=16),
                                                                        axis.text = element_text(size = 16))
```


#### Labeled by Sample Type {.tabset}

```{r}
library(RColorBrewer)

p <- plot_cell_trajectory(podo_cds2, color_by = "condition2", cell_size = 0.5) +
  scale_color_brewer(palette = "Paired") +
  labs(color="", shape="") +
  theme_classic(base_size = 16, base_family = "Roboto") %+replace% theme(legend.position="top",
                                                                        legend.text = element_text(size=16),
                                                                        axis.title = element_text(size=16),
                                                                        axis.text = element_text(size = 16))

p + guides(color = guide_legend(override.aes = list(size = 2)))

``` 



#### Pseudotime plot faceted {.tabset}


##### Color Option 2

```{r, fig.width=8, fig.height=11}

plot_cell_trajectory(podo_cds2, color_by = "condition2", cell_size = 0.2, show_branch_points = F, show_tree = F) +
  scale_color_brewer(palette = "Paired") +
  facet_wrap(~condition2, ncol=2) + 
  theme_classic(base_size = 16, base_family = "Roboto") %+replace% theme(strip.background  = element_blank(),
            panel.grid.major = element_line(colour="white"),
            panel.grid.minor = element_line(colour="white"),
            panel.background = element_blank(), 
            legend.position = "none",
            legend.text = element_text(size=16),
            axis.title = element_text(size=14),
            axis.text = element_text(size = 14),
            strip.text = element_text(size = 14)) 




``` 


### Boxplot


```{r fig.height=8, fig.width=9.75}

pt_plotDF <- pData(podo_cds2)

pt_plotDF <- pt_plotDF %>% filter(cell_type_manual_factor == "Podocyte")

base_color = gg_color_hue(1)
ggplot(data = pt_plotDF, aes(x=condition2,y=Pseudotime,fill=condition2)) +
  geom_boxplot(alpha = 1, show.legend = FALSE, color="grey20") +
  scale_fill_manual(values=c("grey30", "grey50", "#fcaaa4", base_color, "#fcaaa4", base_color, "#fcaaa4", base_color, "#fcaaa4", base_color)) +
  xlab("") +
  theme_bw(base_size = 18, base_family = "ARIAL") %+replace% theme(axis.text.x = element_text(angle = 90, hjust = 1),
                                                                  legend.position="top",
                                                                  legend.text = element_text(size=18),
                                                                  axis.title = element_text(size=18),
                                                                  axis.text = element_text(size = 18))

ggsave("fig2d.png", device = "png", height = 8, width = 9.75, dpi = 600)

```




## Fig s7B

```{r}


p <- plot_cell_trajectory(podo_cds2, color_by = "condition2", cell_size = 0.5) +
  scale_color_brewer(palette = "Paired") +
  labs(color="", shape="") +
  theme_classic(base_size = 16) %+replace% theme(legend.position="top",
                                                                        legend.text = element_text(size=16),
                                                                        axis.title = element_text(size=16),
                                                                        axis.text = element_text(size = 16))

p + guides(color = guide_legend(override.aes = list(size = 2)))

``` 


## Fig s7D


```{r, fig.width=8, fig.height=11}

plot_cell_trajectory(podo_cds2, color_by = "condition2", cell_size = 0.2, show_branch_points = F, show_tree = F) +
  scale_color_brewer(palette = "Paired") +
  facet_wrap(~condition2, ncol=2) + 
  theme_classic(base_size = 16) %+replace% theme(strip.background  = element_blank(),
            panel.grid.major = element_line(colour="white"),
            panel.grid.minor = element_line(colour="white"),
            panel.background = element_blank(), 
            legend.position = "none",
            legend.text = element_text(size=16),
            axis.title = element_text(size=14),
            axis.text = element_text(size = 14),
            strip.text = element_text(size = 14)) 




``` 
