---
title: "Pseudotime analysis - Distal Tubule and Collecting Duct"
output:
  html_document:
    code_folding: hide
    collapsed: no
    highlight: haddock
    keep_md: no
    theme: cerulean
    toc: yes
    toc_float:
      collapsed: yes
    toc_depth: 6
    number_sections: false
params:
    date: !r format(Sys.Date(), "%d %B, %Y")
    client: Goldfinch Bio

---

<!-- Indent the TOC -->
<script>
$(document).ready(function() {
  $items = $('div#TOC li');
  $items.each(function(idx) {
    num_ul = $(this).parentsUntil('#TOC').length;
    $(this).css({'text-indent': num_ul * 10, 'padding-left': 0});
  });

});
</script>

<table width=100%>
<tr>
<td>
<h3>Prepared for `r params$client` by:</h3>
Diamond Age Data Science
</td>
<td width=150></td>
<td width=150 align=right>`r params$date`</td>
</tr>
</table>




```{r setup, include=FALSE}
DOCNAME = "in_vivo_dt_cd"
knitr::opts_chunk$set(autodep        = TRUE,
                      cache          = FALSE,
                      cache.path     = paste0("cache/", DOCNAME, "/"),
                      fig.path       = paste0("cache/", DOCNAME, "/"),
                      cache.comments = FALSE,
                      cache.lazy     = FALSE,
                      echo           = TRUE,
                      error          = FALSE,
                      fig.align      = "center",
                      fig.width      = 10,
                      fig.height     = 8,
                      message        = FALSE,
                      warning        = FALSE)

#Libraries
library(Seurat)
library(parallel)
library(future)
library(here)
library(GSVA)
library(clustree)
library(viridis)
library(knitr)
library(DT)
library(kableExtra)
library(tidyverse)
library(scales)
library(harmony)
library(monocle)
library(slingshot)

source(here("R/utilities.R"))

#Options
nCores <- 12
options(future.globals.maxSize = 8 * 1024^3)
save = TRUE



# Functions



# Directories
baseDir = here()
saveDir = file.path(baseDir,"output", DOCNAME)
if(!dir.exists(saveDir)){
  dir.create(saveDir)
}

dataDir <- "/mnt/cellranger/scRNAseq/output/in_vivo_dt_cd"

```


# Trajectory analysis of Distal Tubule and Collecting Duct

In this analysis I combined the distal tubule and collecting duct cells from all of the in vivo datasets. I also included cells from adult and fetal kidney.

I batch corrected the data using the Harmony package. This allowed for clustering of the adult and kidney cells with the transplants.

```{r}

if(file.exists(file.path(dataDir, "seur.RDS"))){
  seur <- read_rds(file.path(dataDir, "seur.RDS"))
}else{
  
  conditions <- c(  "Day_10_2wk_transplant",
                  "Day_10_4wk_transplant",
                  "Day_12_2wk_transplant",
                  "Day_12_4wk_transplant",
                  "Day_14_2wk_transplant",
                  "Day_14_4wk_transplant", 
                  "human_qc")
  names(conditions) <- conditions
  
  
  select_distal_tubule <- function(c){
    temp_seur <- read_rds(here("output", c, paste0(c, ".RDS")))
    DefaultAssay(temp_seur) <- "RNA"
    # temp_seur$cell_type_manual_podo <- ifelse(temp_seur$cell_type_manual == "Podocyte Precursor", "Podocyte", temp_seur$cell_type_manual)
    # temp_seur$cell_rep <- paste0(temp_seur$orig.ident, "_", temp_seur$cell_type_manual_podo)
    podos <- subset(temp_seur, cell_type_manual %in% c("Distal Tubule", "Collecting Duct"))
    Idents(podos) <- "orig.ident"
    return(podos)
  }
  
  seur_list <- mclapply(conditions, select_distal_tubule, mc.cores = 8)

  
  ### removed seur_list[[2]] had no DT or CD cells
  seur <- merge(seur_list[[1]], c(seur_list[[3]], seur_list[[4]], seur_list[[5]], seur_list[[6]], seur_list[[7]]))
  seur <- subset(seur, orig.ident != "BTN-PH" & orig.ident != "BTN-YY")
  
  
  ########### Harmony integration
  
  plan(strategy="multiprocess", workers = 8)
  seur <- NormalizeData(seur)
  seur <- ScaleData(seur, vars.to.regress = c("nCount_RNA", "nFeature_RNA"))
  seur <- FindVariableFeatures(seur, selection.method = "vst", nfeatures = 2000)
  seur <- RunPCA(seur, npcs=100)
  # ElbowPlot(seur, ndims = 100)
  
  seur$condition <- ifelse(grepl("_rep", seur$orig.ident), str_split(seur$orig.ident, "_rep", simplify = T)[,1], seur$orig.ident)
  seur$condition2 <- ifelse(grepl("Adult",seur$condition), "AdultKidney",
                                  ifelse(grepl("Fetal", seur$condition), "FetalKidney", seur$condition))
  
  seur <- RunUMAP(seur, reduction = "pca", dims = 1:100)
  
  plan(strategy = "sequential")
  
  seur <- RunHarmony(seur, group.by.vars = "condition2")
  seur <- RunUMAP(seur, reduction = "harmony", reduction.name = "umap_harmony", reduction.key = "umap_harmony_", dims = 1:50)
  seur <- FindNeighbors(seur, dims = 1:50, reduction = "harmony")
  seur <- FindClusters(seur, resolution = c(0.05,seq(0,1,0.1)))
  
  saveDat(seur)
  
  
  
  
 
}


condition_levels <- c("AdultKidney", "FetalKidney", "D10-2wk", "D10-4wk", "D12-2wk", "D12-4wk", "D14-2wk", "D14-4wk")

```






## Cluster Markers

```{r}


if(file.exists(file.path(dataDir, "markers.RDS"))){
  markers <- read_rds(file.path(dataDir, "markers.RDS"))
}else{
  
plan(strategy = "multiprocess", workers = 8)

Idents(seur) <- "RNA_snn_res.0.9"
markers <- FindAllMarkers(seur)
saveDat(markers)
plan(strategy = "sequential")
}

cd_markers <- c("AQP1", "AQP2", "AQP3","AQP5", "ATP1A1", "ATP1B1", "CALB1", "NKCC1", "HMX2", "SLC4A1", "CLDN3", "CLDN4", "FOXI1")

# top5 <- markers %>% group_by(cluster) %>% top_n(5, avg_logFC)

# clusters <- dt_markers %>% filter(gene %in% c("AQP2", "CALB1")) %>% filter(avg_logFC > 0)
clusters <- markers %>% filter(gene %in% cd_markers) %>% 
  filter(avg_logFC > 0)
# cd_clusters <- unique(clusters$cluster[duplicated(clusters$cluster)])
cd_clusters <- names(table(clusters$cluster))[table(clusters$cluster) >= 3]

# cd_clusters <- c(1,13,3,9, 6)



```

Chose clusters with increased expression of 3 or more collecting duct marker genes.



## Collecting Duct Clusters

```{r}

# cd_clusters <- c(1,2,5,13,22)

seur$cell_type_manual <- ifelse(seur$RNA_snn_res.0.9 %in% cd_clusters, "Collecting Duct-like", "Distal Tubule")

meta <- seur@meta.data %>% select(cell_type_manual) %>% dplyr::rename(cell_type_manual_3 = cell_type_manual)

seur <- AddMetaData(seur, meta)

seur$cell_type_manual_3 <- ifelse(is.na(seur$cell_type_manual_3), seur$cell_type_manual, seur$cell_type_manual_3)

temp_meta <- tibble(cell_name = rownames(seur@meta.data),
                    ident = seur$cell_type_manual_3)

collapse <- c("S-Shaped Body", "Comma-Shaped Body", "Renal Vessicle", "Cap Mesenchyme")

temp_meta <- temp_meta %>% mutate(cell_type_simple_3 = ifelse(ident %in% collapse, "Nephron Progenitor", ident)) %>% 
  select(cell_name, cell_type_simple_3) %>% column_to_rownames("cell_name") #%>% mutate(cell_type_simple_2 = factor(cell_type_simple_2, levels = cell_levels_collapse))

seur <- AddMetaData(seur, temp_meta)


seur$cell_type_simple_3 <- factor(seur$cell_type_simple_3, levels = c("Distal Tubule", "Collecting Duct-like"))

```


I chose cluster(s) **`r cd_clusters`** for collecting duct






Manual Cell Type Proportions w/ Collecting Duct
-----------------------------------------------

### UMAP {.tabset}

#### Cell types

```{r fig.width=8}
cell_colors <- c(colors["Distal Tubule"],"firebrick")


figs8a <- DimPlot(seur, group.by = "cell_type_simple_3", reduction = "umap_harmony") + 
  scale_color_manual(values=cell_colors) +
  theme_classic(base_size = 18, base_family = "DejaVu Sans") %+replace% theme(legend.position="top",
                                                                         legend.text = element_text(size=18),
                                                                         axis.title = element_text(size=18),
                                                                         axis.text = element_text(size = 18))
figs8a

```



#### Collecting Duct Markers

```{r, fig.width=10, fig.height=10}
# FeaturePlot(seur, features=c("AQP2", "CALB1", "CLDN3", "CDH1"), ncol = 2, reduction = "umap_harmony") +
#   theme_classic(base_size = 18, base_family = "DejaVu Sans") %+replace% theme(
#     legend.text = element_text(size=14)
#   )


plot_list <-
  lapply(c("AQP2", "CALB1", "CLDN3", "CDH1"), function(f){
    p <- FeaturePlot(seur, features=f, reduction = "umap_harmony", pt.size = 0.1) +
    theme_classic(base_size = 18, base_family = "DejaVu Sans") %+replace% theme(plot.title = element_text(size=18),
                                                                         legend.text = element_text(size=18),
                                                                         axis.title = element_text(size=18),
                                                                         axis.text = element_text(size = 18)
    )
  }
)

figs8b <- cowplot::plot_grid(plotlist = plot_list, ncol=2)

figs8b

```


```{r fig.height=4, fig.width=8}


figs8c <- DotPlot(seur, features = c("AQP2", "CALB1", "CDH1", "CLDN3"), group.by = "cell_type_simple_3", scale = T, dot.scale = 15) +
  ylab("") +
  xlab("") +
  theme_classic(base_size = 14, base_family = "DejaVu Sans") %+replace% theme(
    legend.text = element_text(size=14),
    legend.title = element_text(size = 14),
    axis.title = element_text(size=14),
    axis.text = element_text(size = 18)
  )

figs8c
```







### Trajectory {.tabset}

Genes that vary most between clusters


```{r, fig.height=6}
if(file.exists(file.path(dataDir, "dt_cds2.RDS"))){
  dt_cds2 <- read_rds(file.path(dataDir, "dt_cds2.RDS"))
}else{
  
  dt_cds2 <- as.CellDataSet(seur, assay = "RNA", reduction = "umap_harmony")
  
  dt_cds2 <- detectGenes(dt_cds2, min_expr = 0.1)
  expressed_genes <- row.names(subset(fData(dt_cds2),
      num_cells_expressed >= 10))
  
  dt_cds2 <- estimateSizeFactors(dt_cds2)
  dt_cds2 <- estimateDispersions(dt_cds2)
  
  
  dt_cds2$batch <- ifelse(dt_cds2$condition2 %in% c("AdultKidney", "FetalKidney"), 1, 2)
  
  clustering_DEG_genes <- differentialGeneTest(dt_cds2[expressed_genes,],
                                               fullModelFormulaStr = "~RNA_snn_res.0.9 + batch",
                                               reducedModelFormulaStr = "~batch",
                                               cores = 12)
  
  ordering_genes <-
      row.names(clustering_DEG_genes)[order(clustering_DEG_genes$qval)][1:1000]
  
  dt_cds2 <-
      setOrderingFilter(dt_cds2,
          ordering_genes = ordering_genes)
  
  dt_cds2 <-
      reduceDimension(dt_cds2, method = 'DDRTree', verbose = T)
  
  dt_cds2 <-
      orderCells(dt_cds2, num_paths = 1, reverse = T)
  
  dt_cds2$sample_name <- NULL
  saveDat(dt_cds2)
}

```

```{r}
dt_cds2$cell_type_simple_3 <- factor(dt_cds2$cell_type_simple_3, levels = c("Distal Tubule", "Collecting Duct-like"))


```



#### Labeled by cell type

```{r}
p <- plot_cell_trajectory(dt_cds2, color_by = "cell_type_simple_3", cell_size = 0.2)  +
  scale_color_manual(values = cell_colors, name="") +
  labs(color="", shape="") +
  theme_classic(base_size = 18, base_family = "DejaVu Sans") %+replace% theme(legend.position="top",
                                                                        legend.text = element_text(size=18),
                                                                        axis.title = element_text(size=18),
                                                                        axis.text = element_text(size = 18))



figs8d <- p + guides(color = guide_legend(override.aes = list(size = 3)))

figs8d

```

#### Labeled by cell state

```{r}
plot_cell_trajectory(dt_cds2, color_by = "State", cell_size = 0.5)
```

#### Labeled by Pseudotime

```{r}

figs8f <- plot_cell_trajectory(dt_cds2, color_by = "Pseudotime", cell_size = 0.5) +
  scale_color_viridis_c(name="") +
  labs(color="", shape="") +
  guides(color = guide_colourbar(barwidth = 10)) +
  theme_classic(base_size = 18, base_family = "DejaVu Sans") %+replace% theme(legend.position="top",
                                                                        legend.text = element_text(size=16),
                                                                        axis.title = element_text(size=18),
                                                                        axis.text = element_text(size = 18))
figs8f

```





#### Labeled by Sample Type {.tabset}

```{r}
conditionDF <- pData(dt_cds2) %>% mutate(condition2 = gsub("day_", "D", condition2),
                            condition2 = gsub("_2wk_transplant", "-2wk", condition2),
                            condition2 = gsub("_4wk_transplant", "-4wk", condition2),
                            # condition2 = gsub("Kidney", " Kidney", condition2),
                            condition2 = factor(condition2, levels = condition_levels))
dt_cds2$condition2 <- conditionDF$condition2
```



##### Color option 1

```{r}

p <- plot_cell_trajectory(dt_cds2, color_by = "condition2", cell_size = 0.25) +
  scale_color_hue(h = c(200, 400), drop=F) +
  labs(color="", shape="") +
  theme_classic(base_size = 18, base_family = "DejaVu Sans") %+replace% theme(legend.position="top",
                                                                        legend.text = element_text(size=16),
                                                                        axis.title = element_text(size=18),
                                                                        axis.text = element_text(size = 18),
                                                                        legend.spacing.x = unit(0.75, 'cm'))


p + guides(color = guide_legend(override.aes = list(size = 3)))



```


##### Color option 2

```{r}
p <- plot_cell_trajectory(dt_cds2, color_by = "condition2", cell_size = 0.25) +
  scale_color_viridis_d(name="", option = "inferno", drop=FALSE) +
  labs(color="", shape="") +
  theme_classic(base_size = 18, base_family = "DejaVu Sans") %+replace% theme(legend.position="top",
                                                                        legend.text = element_text(size=18),
                                                                        axis.title = element_text(size=18),
                                                                        axis.text = element_text(size = 18))

p + guides(color = guide_legend(override.aes = list(size = 2)))

```


##### Color option 3

```{r}
library(RColorBrewer)

p <- plot_cell_trajectory(dt_cds2, color_by = "condition2", cell_size = 0.5) +
  scale_color_brewer(palette = "Paired", drop=FALSE) +
  labs(color="", shape="") +
  theme_classic(base_size = 18, base_family = "DejaVu Sans") %+replace% theme(legend.position="top",
                                                                        legend.text = element_text(size=16),
                                                                        axis.title = element_text(size=18),
                                                                        axis.text = element_text(size = 18))

figs8e <- p + guides(color = guide_legend(override.aes = list(size = 2)))

figs8e

```



#### Labeled by Dataset

```{r}
plot_cell_trajectory(dt_cds2, color_by = "orig.ident", cell_size = 0.5)
```


##



```{r, fig.height=30, fig.width=24}



figl <- cowplot::plot_grid(figs8a, NULL, figs8b, NULL, figs8c, ncol = 1, rel_heights = c(1,0.1,1,0.1,0.5), rel_widths = c(1,0.1,1,0.1,0.5))
figr <- cowplot::plot_grid(figs8d, figs8e, figs8f, ncol = 1)

figs8 <- cowplot::plot_grid(figl, NULL, figr, ncol=3, rel_widths = c(1,0.1,1))
figs8

ggsave("figs8.png", plot=figs8, device = "png")

```









