---
title: "QC Human data"
output:
  html_document:
    code_folding: hide
    collapsed: no
    highlight: haddock
    keep_md: no
    theme: cerulean
    toc: yes
    toc_float:
      collapsed: yes
    toc_depth: 6
    number_sections: false
params:
    date: !r format(Sys.Date(), "%d %B, %Y")
    client: Goldfinch Bio

---

<!-- Indent the TOC -->
<script>
$(document).ready(function() {
  $items = $('div#TOC li');
  $items.each(function(idx) {
    num_ul = $(this).parentsUntil('#TOC').length;
    $(this).css({'text-indent': num_ul * 10, 'padding-left': 0});
  });

});
</script>

<table width=100%>
<tr>
<td>
<h3>Prepared for `r params$client` by:</h3>
Diamond Age Data Science
</td>
<td width=150></td>
<td width=150 align=right>`r params$date`</td>
</tr>
</table>




```{r setup, include=FALSE}
DOCNAME = "human_qc"
knitr::opts_chunk$set(autodep        = TRUE,
                      cache          = TRUE,
                      cache.path     = paste0("cache/", DOCNAME, "/"),
                      fig.path       = paste0("cache/", DOCNAME, "/"),
                      cache.comments = FALSE,
                      cache.lazy     = FALSE,
                      echo           = TRUE,
                      error          = FALSE,
                      fig.align      = "center",
                      fig.width      = 10,
                      fig.height     = 8,
                      message        = FALSE,
                      warning        = FALSE)

#Libraries
library(Seurat)
library(parallel)
library(future)
library(here)
library(GSVA)
library(clustree)
library(viridis)
library(knitr)
library(DT)
library(kableExtra)
library(tidyverse)
library(monocle3)

source(here("R/utilities.R"))

#Options
nCores <- 18
options(future.globals.maxSize = 8 * 1024^3)
save = TRUE



# Functions



# Directories
baseDir = here()
saveDir = file.path(baseDir,"output", DOCNAME)
if(!dir.exists(saveDir)){
  dir.create(saveDir)
}



```




Load counts
-----------

```{r process-counts}


# data_sets <- list.files(here("data/human"))
# data_sets <- data_sets[!grepl(".xlsx", data_sets)]

names <- c("Adult_Kidney_GSE114156", "Fetal_Kidney_115days", "Fetal_Kidney_110days", "Fetal_Kidney_105days", "Adult_Kidney_Nehprectomy")

Adult_Kidney_GSE114156 <- read_tsv(here("data/human/GSM2935201_FetalKidney_Run1751_105days.txt")) %>% column_to_rownames(var = "GENE")
Adult_Kidney_Nehprectomy <- read_csv(here("data/human/SingleCell_Tumor_nephrectomy_Raw_datamatrix.csv")) %>% filter(!duplicated(GENE)) %>% column_to_rownames(var = "GENE")
Fetal_Kidney_115days <- read_tsv(here("data/human/GSM2935199_FetalKidney_Run1824_115days.txt"))%>% column_to_rownames(var = "GENE")
Fetal_Kidney_110days <- read_tsv(here("data/human/GSM2935200_FetalKidney_Run1785_110days.txt"))%>% column_to_rownames(var = "GENE")
Fetal_Kidney_105days <- read_tsv(here("data/human/GSM2935201_FetalKidney_Run1751_105days.txt"))%>% column_to_rownames(var = "GENE")


counts <- list(Adult_Kidney_GSE114156, Fetal_Kidney_115days, Fetal_Kidney_110days, Fetal_Kidney_105days, Adult_Kidney_Nehprectomy)
names(counts) <- names



# 
# 
# 
# 
# 
# counts <- mclapply(1:length(data_sets), function(x){
#     counts <- read_tsv(here("data/human", data_sets[x]))
#     colnames(counts) <- paste0(colnames(counts), "-", names(data_sets)[x])
#     return(counts)
#     }, mc.cores = nCores)



seur.list <- mclapply(1:length(counts), function(x) CreateSeuratObject(counts[[x]],min.cells = 10, project = names(counts)[x],
                                                                       meta.data = data.frame(row.names = colnames(counts[[x]]),
                                                                                              orig.project = rep(names(counts)[x], ncol(counts[[x]])))))

seur <- merge(seur.list[[1]], y=c(seur.list[[1]], seur.list[[2]], seur.list[[3]], seur.list[[4]], seur.list[[5]]))
seur <- PercentageFeatureSet(seur, pattern = "MT-", col.name = "percent.mito")


```



Create Seurat Object
--------------------



## QC Plots 

### UMI counts vs Number of Genes


```{r seur_scatter}
# FeatureScatter(seur, feature1 = "nCount_RNA", feature2 = "nFeature_RNA", group.by = "species") + ggtitle("All Cells")

ggplot(data = seur@meta.data, aes(x=nCount_RNA, nFeature_RNA))+
  geom_jitter(aes(color = "red"), size=0.5)+
  facet_wrap(~orig.ident)

```



### UMI counts vs Percent Mito

```{r seur_scatter_mito}

ggplot(data = seur@meta.data, aes(x=nCount_RNA, y=percent.mito))+
  geom_jitter(aes(color= "red"), size=0.5) +
  facet_wrap(~orig.ident) + NoLegend()

```


### Violin Plots
QC violin plots grouped by orig ident

```{r qc_violin, fig.height=4}
VlnPlot(seur, features = c("nFeature_RNA", "nCount_RNA", "percent.mito"), pt.size = 0, group.by = "orig.ident")

```

### QC Box Plots

```{r}
melt <- reshape2::melt(seur@meta.data %>% select(nFeature_RNA, nCount_RNA, percent.mito, orig.ident), id.vars=c("orig.ident"))
# melt$orig.ident <- factor(melt$orig.ident)

ggplot(data = melt, aes(x=orig.ident, y=value, fill=orig.ident))+
  geom_boxplot() +
  facet_wrap(~variable, scales = "free") +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```



## Filter Cells


```{r filter_cells}

seur <- subset(seur, nFeature_RNA > 100 & nCount_RNA > 100 & nCount_RNA < 10000 & percent.mito < 20)

```

### Post-QC Violin Plots

```{r post_qc_vln}

VlnPlot(seur, features = c("nFeature_RNA", "nCount_RNA", "percent.mito"), pt.size = 0, group.by = "orig.ident")



```

### Post-QC Box Plots

```{r post_qc_box}

seur$orig.ident <- factor(seur$orig.ident)

seur_box_plot(object = seur,
              features = c("nFeature_RNA", "nCount_RNA", "percent.mito"),
              group = "orig.ident")


```

### Cell Numbers

This graph shows the number of human cells in each orig.ident after filtering

```{r post_qc_numb, fig.height=4, fig.width=4}

plotDF <- seur@meta.data
# plotDF %>% group_by(orig.ident) %>% summarize(count = n())

ggplot(data = plotDF, aes(x=orig.ident, fill= orig.ident))+
  geom_bar() +
  theme(legend.position = "none")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```



```{r, cache=FALSE}
if(file.exists(file.path(saveDir, paste0(DOCNAME, ".RDS")))){
  seur <- read_rds(file.path(saveDir, paste0(DOCNAME, ".RDS")))
  do_eval <- FALSE
}else{
  do_eval <- TRUE
}

```


Integrate
=========

```{r prep-seurList, eval=do_eval, cache=!do_eval, results="hide"}
seurList <- SplitObject(seur, split.by = "orig.ident")

prepSeur <- function(x){
                        temp <- seurList[[x]]
                        temp <- NormalizeData(temp, verbose = FALSE)
                        temp <- FindVariableFeatures(temp, selection.method = "vst", nfeatures = 2000, verbose = FALSE)
}

seurList <- mcmapply(prepSeur, names(seurList), mc.cores = nCores)

```


```{r integrate-seurat, eval=do_eval, cache=!do_eval, results="hide"}
plan(strategy = "multiprocess", workers=8)
anchors <- FindIntegrationAnchors(seurList, anchor.features = 2000, dims = 1:50, verbose = TRUE)
plan(strategy = "sequential")
gc()
```



```{r int-seur-2, eval=do_eval, cache=!do_eval, results="hide"}

plan(strategy = "multiprocess", workers=12)
seur <- IntegrateData(anchors, dims = 1:50, verbose = TRUE)
plan(strategy = "sequential")
saveDat(seur, filename = DOCNAME)
rm(anchors, seurList, tempSeur)
gc()
```


Normalized and Scale Data
-------------------------

```{r scale-int, eval=do_eval, cache=!do_eval}


plan(strategy = "multiprocess", workers=12)
seur <- ScaleData(object = seur,
                  verbose = TRUE)

plan(strategy = "sequential")



saveDat(seur, filename = DOCNAME)
```


```{r}
seur <- RunPCA(seur, npcs = 75)
seur <- RunUMAP(seur, dims = 1:75)

saveDat(seur, filename = DOCNAME)
```




```{r}
DimPlot(seur, group.by = "orig.ident")
```

```{r}
FeaturePlot(seur, features = c("NPHS1", "NPHS2"))
```

```{r, eval=do_eval, cache=!do_eval}
seur <- FindNeighbors(seur, dims = 1:75)
seur <- FindClusters(seur, resolution = 0.5)

saveDat(seur, filename = DOCNAME)
```


```{r}
DimPlot(seur, group.by = "integrated_snn_res.0.5", label = T)
```



```{r, eval=do_eval, cache=!do_eval}
seur$cell_type_manual <- ifelse(seur$integrated_snn_res.0.5 == 4, "Podocyte", "other")

saveDat(seur, filename = DOCNAME)

```


```{r}
podos <- subset(seur, integrated_snn_res.0.5 == 4)

Idents(podos) <- "orig.ident"

podos <- ScaleData(podos, assay = "RNA")

exprs <- AverageExpression(podos, assays = "RNA")

saveDat(exprs)
```





```{r post_qc_numb2, fig.height=6, fig.width=4}

plotDF <- podos@meta.data
# plotDF %>% group_by(replicate) %>% summarize(count = n())

ggplot(data = plotDF, aes(x=orig.ident, fill= orig.ident))+
  geom_bar() +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

cell_num <- plotDF %>% group_by(orig.ident) %>% summarise(count = n())
cell_num %>% kable() %>% kable_styling()


num <- sum(cell_num$count)
```


```{r}
VlnPlot(seur, features = c("GATA3", "MAL", "CDH1", "AQP2"), assay = "RNA", pt.size = 0.1, ncol=2)
```

```{r}
FeaturePlot(seur, features = c("GATA3", "MAL", "CDH1", "AQP2"), min.cutoff = "q9")


```


Distal Tubule/Collecting Duct in cluster 3

```{r}
seur$cell_type_manual <- ifelse(seur$integrated_snn_res.0.5 == 3, "Distal Tubule", seur$cell_type_manual)
table(seur$cell_type_manual)

seur$cell_type_simple_2 <- seur$cell_type_manual
```


## Subset Distal Tubule Cells

```{r distal-tubule-recluster, results="hide"}


if(file.exists(file.path(saveDir, "dt_seur.RDS"))){
  dt_seur <- read_rds(file.path(saveDir, "dt_seur.RDS"))
}else{
  DefaultAssay(seur) <- "RNA"
  dt_seur <- subset(seur, cell_type_manual == "Distal Tubule")
  
  seurList <- SplitObject(dt_seur, split.by = "orig.ident")
  seurList <- mcmapply(prepSeur, names(seurList), mc.cores = 8)
  
  k.filter <- min(200, min(sapply(seurList, ncol)))
  anchors <- FindIntegrationAnchors(seurList, anchor.features = 2000, dims = 1:30, verbose = TRUE, k.filter = k.filter)
  dt_seur <- IntegrateData(anchors, dims = 1:50, verbose = TRUE)
  
  dt_seur <- ScaleData(dt_seur, vars.to.regress = c("orig.ident", "nCount_RNA", "percent.mito", "S.Score", "G2M.Score"), verbose = TRUE)
  
  dt_seur <- RunPCA(dt_seur, npcs = 50, verbose=F)
  dt_seur <- RunUMAP(dt_seur, dims = 1:50)
  
  dt_seur <- FindNeighbors(dt_seur, dims = 1:50)
  res = seq(0,2,0.1)
  dt_seur <- FindClusters(dt_seur, resolution = res)
  # dt_seur$RNA_snn_res.3 <- factor(dt_seur$RNA_snn_res.3, levels = min(as.numeric(levels(dt_seur$RNA_snn_res.3))):max(as.numeric(levels(dt_seur$RNA_snn_res.3))))
  
  DefaultAssay(dt_seur) <- "RNA"
  
  saveDat(dt_seur)
}
```


## UMAP {.tabset}

### Grouped by sample

```{r, fig.height=6, fig.width=10}
DimPlot(dt_seur, group.by = "orig.ident")
```


### Grouped by cluster

```{r, fig.height=6, fig.width=8}
DimPlot(dt_seur, group.by="integrated_snn_res.1.5", label = T)
```

### Collecting Duct Markers

```{r, fig.height=10}
FeaturePlot(dt_seur, features = c("AQP2", "CALB1", "CDH1", "KRT8"), min.cutoff = "q9", ncol = 2)

```

## Marker Gene Violin Plots {.tabset}

### Collecting Duct

```{r, fig.height=6}
VlnPlot(dt_seur, features = c("AQP2", "CALB1", "CDH1", "KRT8"), group.by = "integrated_snn_res.1.5", ncol = 2, pt.size = 0.1)

```


### Distal Tubule

```{r, fig.height=6}
VlnPlot(dt_seur, features = vln_markers_list[["Distal Tubule"]]$Gene, group.by = "integrated_snn_res.1.5", ncol = 2, pt.size = 0.1)
```

## Cluster Markers

```{r}


if(file.exists(file.path(saveDir, "dt_markers.RDS"))){
  dt_markers <- read_rds(file.path(saveDir, "dt_markers.RDS"))
}else{
  
plan(strategy = "multiprocess", workers = 8)

Idents(dt_seur) <- "integrated_snn_res.1.5"
dt_markers <- FindAllMarkers(dt_seur)
saveDat(dt_markers)
plan(strategy = "sequential")
}

cd_markers <- c("AQP1", "AQP2", "AQP3","AQP5", "ATP1A1", "ATP1B1", "CALB1", "NKCC1", "HMX2", "SLC4A1", "CLDN3", "CLDN4", "FOXI1")

top5 <- dt_markers %>% group_by(cluster) %>% top_n(5, avg_logFC)

# clusters <- dt_markers %>% filter(gene %in% c("AQP2", "CALB1")) %>% filter(avg_logFC > 0)
clusters <- dt_markers %>% filter(gene %in% cd_markers) %>% 
  filter(avg_logFC > 0)
# cd_clusters <- unique(clusters$cluster[duplicated(clusters$cluster)])
cd_clusters <- names(table(clusters$cluster))[table(clusters$cluster) >= 3]





```

Chose clusters with increased expression of 3 or more collecting duct marker genes.

`r kable(as.data.frame(cd_markers) %>% rename("Markers"=cd_markers))`


```{r, fig.height=12, fig.width=12}
DoHeatmap(dt_seur, top5$gene, label = F)
```


## Collecting Duct Clusters

```{r}

# cd_clusters <- c(1,2,5,13,22)

dt_seur$cell_type_manual <- ifelse(dt_seur$integrated_snn_res.1.5 %in% cd_clusters, "Collecting Duct", "Distal Tubule")

meta <- dt_seur@meta.data %>% select(cell_type_manual) %>% dplyr::rename(cell_type_manual_2 = cell_type_manual)

seur <- AddMetaData(seur, meta)

seur$cell_type_manual_2 <- ifelse(is.na(seur$cell_type_manual_2), seur$cell_type_manual, seur$cell_type_manual_2)

temp_meta <- tibble(cell_name = rownames(seur@meta.data),
                    ident = seur$cell_type_manual_2)

collapse <- c("S-Shaped Body", "Comma-Shaped Body", "Renal Vessicle", "Cap Mesenchyme")

temp_meta <- temp_meta %>% mutate(cell_type_simple_2 = ifelse(ident %in% collapse, "Nephron Progenitor", ident)) %>% 
  select(cell_name, cell_type_simple_2) %>% column_to_rownames("cell_name") #%>% mutate(cell_type_simple_2 = factor(cell_type_simple_2, levels = cell_levels_collapse))

seur <- AddMetaData(seur, temp_meta)


seur$cell_type_simple_2 <- factor(seur$cell_type_simple_2, levels = cell_levels_collapse)
```


I chose cluster(s) **`r cd_clusters`** for collecting duct



Manual Cell Type Proportions w/ Collecting Duct
-----------------------------------------------

### UMAP {.tabset}

#### All Cell types

```{r}
DimPlot(seur, group.by = "cell_type_manual_2")
```


#### Simplified Cell Types

```{r}
DimPlot(seur, group.by = "cell_type_simple_2", cols = colors)
```


#### Collecting Duct Markers

```{r, fig.height=10, fig.width=10}
FeaturePlot(seur, features=c("AQP2", "CALB1", "CDH1", "KRT8"), ncol = 2)
```





```{r}
saveDat(seur, DOCNAME)
```








