---
title: "in vivo Cell Types"
output:
  html_document:
    code_folding: hide
    collapsed: no
    highlight: haddock
    keep_md: no
    theme: cerulean
    toc: yes
    toc_float:
      collapsed: yes
    toc_depth: 6
    number_sections: false
params:
    date: !r format(Sys.Date(), "%d %B, %Y")
    client: Goldfinch Bio

---

<!-- Indent the TOC -->
<script>
$(document).ready(function() {
  $items = $('div#TOC li');
  $items.each(function(idx) {
    num_ul = $(this).parentsUntil('#TOC').length;
    $(this).css({'text-indent': num_ul * 10, 'padding-left': 0});
  });

});
</script>

<table width=100%>
<tr>
<td>
<h3>Prepared for `r params$client` by:</h3>
Diamond Age Data Science
</td>
<td width=150></td>
<td width=150 align=right>`r params$date`</td>
</tr>
</table>



```{r setup, include=FALSE}
DOCNAME = "in_vivo_cell_types_12-2019"
knitr::opts_chunk$set(autodep        = TRUE, 
                      cache          = FALSE,
                      cache.path     = paste0("cache/", DOCNAME, "/"),
                      fig.path       = paste0("cache/", DOCNAME, "/"),
                      cache.comments = FALSE,
                      cache.lazy     = FALSE,
                      echo           = TRUE,
                      error          = FALSE,
                      fig.align      = "center",
                      fig.width      = 10,
                      fig.height     = 8,
                      dpi            = 600,
                      message        = FALSE,
                      warning        = FALSE)

#Libraries
library(here)
library(viridis)
library(knitr)
library(DT)
library(kableExtra)
library(philentropy)
library(tidyverse)
library(Seurat)


source(here("R/utilities.R"))
rename = dplyr::rename


#Options
save = TRUE



# Directories
baseDir = here()
saveDir = file.path(baseDir,"output", DOCNAME)
if(!dir.exists(saveDir)){
  dir.create(saveDir, recursive = T)
}



```





```{r}
samples <- c("Day_10_2wk", "Day_10_4wk", "Day_12_2wk", "Day_12_4wk", "Day_14_2wk", "Day_14_4wk")

cell_type_list <- lapply(samples, function(s){
  csv <- list.files(path = here("output", paste0(s, "_transplant")), pattern = "props.csv")
  df <- read_csv(here("output", paste0(s, "_transplant"),csv))
  df$Condition <- rep(s, nrow(df))
  df$Condition <- gsub("_", " ", df$Condition)
  df <- df %>% rename("Sample"=orig.ident)
  df$Sample <- str_split(df$Sample, "_", simplify = T)[,5]
  df$Sample <- gsub("rep", "Rep ", df$Sample)
  return(df)
})

names(cell_type_list) <- samples

cell_type_df <- do.call(rbind, cell_type_list)
cell_type_df <- cell_type_df %>% mutate(Condition = gsub("Day ", "D", Condition),
                                        Condition = gsub(" 2wk", "-2wk", Condition),
                                        Condition = gsub(" 4wk", "-4wk", Condition))

condition_levels <- c("D10-2wk", "D10-4wk", "D12-2wk", "D12-4wk", "D14-2wk", "D14-4wk")

cell_type_df <- cell_type_df %>% mutate(Condition = factor(Condition, levels = condition_levels))


cell_type_df$`Cell Type` <- gsub("Neural", "Neuronal", cell_type_df$`Cell Type`,fixed = TRUE)


```

```{r}
write_csv(cell_type_df, file.path(saveDir, "cell_type_counts.csv"))

cell_types <- as_tibble(unique(cell_type_df$`Cell Type`)[order(unique(cell_type_df$`Cell Type`))])

write_csv(cell_types, file.path(saveDir, "cell_types.csv"))
```



```{r}

cell_type_df$`Cell Type` <- factor(cell_type_df$`Cell Type`, levels = cell_levels)


```



```{r}
samples <- c("Day_10_2wk", "Day_10_4wk", "Day_12_2wk", "Day_12_4wk", "Day_14_2wk", "Day_14_4wk")

metadata_list <- lapply(samples, function(s){
  csv <- list.files(path = here("output", paste0(s, "_transplant")), pattern = "_meta.csv")
  df <- read_csv(here("output", paste0(s, "_transplant"),csv))
  df <- df %>% select(orig.ident, cell_type_manual, cell_type_manual_2, cell_type_simple, cell_type_simple_2)
  df$Condition <- rep(s, nrow(df))
  df$Condition <- gsub("_", " ", df$Condition)
  df <- df %>% rename("Sample"=orig.ident)
  df$Sample <- str_split(df$Sample, "_", simplify = T)[,5]
  df$Sample <- gsub("rep", "Rep ", df$Sample)
  return(df)
})

names(metadata_list) <- samples

metadata_df <- do.call(rbind, metadata_list)
metadata_df <- metadata_df %>% mutate(Condition = gsub("Day ", "D", Condition),
                                        Condition = gsub(" 2wk", "-2wk", Condition),
                                        Condition = gsub(" 4wk", "-4wk", Condition))

condition_levels <- c("D10-2wk", "D10-4wk", "D12-2wk", "D12-4wk", "D14-2wk", "D14-4wk")

metadata_df <- metadata_df %>% mutate(Condition = factor(Condition, levels = condition_levels))

metadata_df$cell_type_manual <- gsub("Neural", "Neuronal", metadata_df$cell_type_manual, fixed = TRUE)
metadata_df$cell_type_simple <- gsub("Neural", "Neuronal", metadata_df$cell_type_simple, fixed = TRUE)

cell_type_counts <- metadata_df %>% group_by(Condition, cell_type_simple, Sample) %>% summarise(count = n())
# cell_type_counts

cell_type_mean <- cell_type_counts %>% group_by(cell_type_simple, Condition) %>% summarise(mean = mean(count))
# cell_type_mean


cell_type_counts_day_week <- metadata_df %>% separate(Condition, into=c("Day", "Week"), remove = F) %>% group_by(cell_type_simple, Day, Week) %>% summarise(count = n())
# cell_type_counts_day_week

cell_type_mean_day <- cell_type_counts_day_week %>% group_by(cell_type_simple, Day) %>% summarise(mean = mean(count))
# cell_type_mean_day

cell_type_mean_week <- cell_type_counts_day_week %>% group_by(cell_type_simple, Week) %>% summarise(mean = mean(count))
# cell_type_mean_week




write_csv(cell_type_counts, file.path(saveDir, "cell_type_counts.csv"))
write_csv(cell_type_mean, file.path(saveDir, "cell_type_median.csv"))
write_csv(cell_type_counts_day_week, file.path(saveDir, "cell_type_counts_day_week.csv"))
write_csv(cell_type_mean_day, file.path(saveDir, "cell_type_mean_day.csv"))
write_csv(cell_type_mean_week, file.path(saveDir, "cell_type_mean_week.csv"))
```




# ***in vivo*** Cell Type Proportions

## All Cell Types


```{r, fig.width=10}

# ggplot(data = cell_type_df, aes(x=`Cell Type`, y = Pct))+
#   geom_bar(stat = "identity") +
#   facet_wrap(~ Condition)


ggplot(cell_type_df %>% group_by(Condition), aes(x=Sample, y=Pct, fill=`Cell Type`)) +
  geom_bar(stat="identity", position = "fill", color = "white") +
  labs(title="", y="Proportion", x="", fill="Cell Type") +
  facet_grid(~ Condition) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  theme_bw(base_size = 14) %+replace% theme(axis.text.x = element_text(angle = 90, hjust = 1))


# p2 + theme_minimal()


```





## Condensed Cell Types

### Figure 1B

```{r, fig.width=11.3, fig.height=9}

collapse <- c("S-Shaped Body", "Comma-Shaped Body", "Renal Vessicle", "Cap Mesenchyme")

plotDF <- cell_type_df %>% mutate(cell_type_character = as.character(`Cell Type`)) %>% 
  mutate(cell_type_character = ifelse(cell_type_character %in% collapse, "Nephron Progenitor", cell_type_character),
         `Cell Type` = factor(cell_type_character, levels = cell_levels_collapse),
         Condition = factor(Condition, levels = c("D10-2wk", "D12-2wk", "D14-2wk", "D10-4wk", "D12-4wk", "D14-4wk")))

plotDF %>% group_by(Condition, `Cell Type`, cell_type_character, Sample) %>% summarize_at(vars(Pct), sum) -> plotDF


ggplot(plotDF %>% group_by(Condition), aes(x=Sample, y=Pct, fill=`Cell Type`)) +
  geom_bar(stat="identity", position = "fill", color = "white") +
  labs(title="", y="Proportion", x="", fill="Cell Type") +
  facet_grid(~ Condition) +
  # scale_fill_manual(values = c(gg_color_hue(6), "grey90", "grey80", "grey70", "grey60", "grey30")) +
  scale_fill_manual(values = colors, drop = FALSE, name="") +
  theme_default(base_size = 16) %+replace% theme(axis.text.x = element_text(angle = 90, hjust = 1),
                                            legend.text = element_text(size=18),
                                             axis.text = element_text(size=18),
                                             strip.text = element_text(size = 16, margin = margin(0.1,0,0.1,0, "cm")),
                                             axis.title = element_text(size = 18))

ggsave("journal_submission/fig2b.png", device = "png", width = 11.3, height = 9, dpi = 600)



# saveDat(plotDF, filename = "condensed_cell_proportions", csv = TRUE)
```




## Facet by week

```{r, fig.height=12, fig.width=6}
collapse <- c("S-Shaped Body", "Comma-Shaped Body", "Renal Vessicle", "Cap Mesenchyme")


plotDF <- cell_type_df %>% mutate(cell_type_character = as.character(`Cell Type`)) %>% 
  mutate(cell_type_character = ifelse(cell_type_character %in% collapse, "Nephron Progenitor", cell_type_character),
         `Cell Type` = factor(cell_type_character, levels = cell_levels_collapse))

plotDF %>% group_by(Condition, `Cell Type`, cell_type_character, Sample) %>% summarize_at(vars(Pct), sum) -> plotDF

plotDF <- plotDF %>% mutate(Day = str_split(Condition, "-", simplify = T)[,1],
                  week = str_split(Condition, "-", simplify = T)[,2])



ggplot(plotDF %>% group_by(Condition), aes(x=Sample, y=Pct, fill=`Cell Type`)) +
  geom_bar(stat="identity", position = "fill", color = "white") +
  labs(title="", y="Proportion", x="", fill="Cell Type") +
  # scale_fill_manual(values = c(gg_color_hue(7), "grey90", "grey70", "grey50", "grey30")) +
  scale_fill_manual(values = colors, drop = FALSE) +
  facet_grid(rows = vars(Day), cols = vars(week), switch = "y") +
  # theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  theme_default() %+replace% theme(axis.text.x = element_text(angle = 90, hjust = 1))

in_vivo_cell_types <- plotDF

```





# Sample Divergence


```{r}


similarityDF <- plotDF %>% mutate(sim_cond = paste0(Day, "_", week, "_", Sample)) %>% mutate(sim_cond = gsub(" ", "", sim_cond))

sim_wide <- reshape2::dcast(similarityDF, sim_cond~`Cell Type`, value.var = "Pct", fill = 0)
sim_wide <- sim_wide %>% column_to_rownames("sim_cond")

sim_dis <- distance(sim_wide, method = "jensen-shannon")

rownames(sim_dis) <- rownames(sim_wide)
colnames(sim_dis) <- rownames(sim_wide)

sim_dis <- as.data.frame(sim_dis) %>% rownames_to_column("Condition") %>% separate(Condition, into = c("Day_x", "Week_x", "Rep_x"), remove = FALSE)
sim_dis <- sim_dis %>% select(Condition, Day_x, Week_x, Rep_x, everything())

melt_sim_dis <- reshape2::melt(sim_dis, id.vars=c("Condition", "Day_x","Week_x", "Rep_x"))

melt_sim_dis <- melt_sim_dis %>% separate(variable, into = c("Day_y", "Week_y", "Rep_y"), remove = FALSE) %>%
  select(Condition, Day_x, Week_x, Rep_x, Day_y, Week_y, Rep_y, everything())

# melt_sim_dis <- melt_sim_dis %>% mutate(comp_x = paste0(Condition, "_", variable),
#                                           comp_y = paste0(variable, "_", Condition))



week_levels <- c(
  paste0("D10-2wk", " vs ", c("D10-2wk", "D12-2wk", "D14-2wk", "D10-4wk", "D12-4wk", "D14-4wk")),
  paste0("D10-4wk", " vs ", c("D10-2wk", "D12-2wk", "D14-2wk", "D10-4wk", "D12-4wk", "D14-4wk")),
  paste0("D12-2wk", " vs ", c("D10-2wk", "D12-2wk", "D14-2wk", "D10-4wk", "D12-4wk", "D14-4wk")),
  paste0("D12-4wk", " vs ", c("D10-2wk", "D12-2wk", "D14-2wk", "D10-4wk", "D12-4wk", "D14-4wk")),
  paste0("D14-2wk", " vs ", c("D10-2wk", "D12-2wk", "D14-2wk", "D10-4wk", "D12-4wk", "D14-4wk")),
  paste0("D14-4wk", " vs ", c("D10-2wk", "D12-2wk", "D14-2wk", "D10-4wk", "D12-4wk", "D14-4wk"))
)


day_levels <- c(
  paste0("D10-2wk", " vs ", c("D10-2wk", "D12-2wk", "D14-2wk", "D10-4wk", "D12-4wk", "D14-4wk")),
  paste0("D12-2wk", " vs ", c("D10-2wk", "D12-2wk", "D14-2wk", "D10-4wk", "D12-4wk", "D14-4wk")),
  paste0("D14-2wk", " vs ", c("D10-2wk", "D12-2wk", "D14-2wk", "D10-4wk", "D12-4wk", "D14-4wk")),
  paste0("D10-4wk", " vs ", c("D10-2wk", "D12-2wk", "D14-2wk", "D10-4wk", "D12-4wk", "D14-4wk")),
  paste0("D12-4wk", " vs ", c("D10-2wk", "D12-2wk", "D14-2wk", "D10-4wk", "D12-4wk", "D14-4wk")),
  paste0("D14-4wk", " vs ", c("D10-2wk", "D12-2wk", "D14-2wk", "D10-4wk", "D12-4wk", "D14-4wk"))
)
```



## Distance Between Days (within weeks)



```{r fig.width=10, fig.height=5}
### Fig S5B

ymin = 0
ymax = 0.3


plotDF2 <- melt_sim_dis %>% mutate(Comparison = paste0(Day_x,"-", Week_x, " vs ", Day_y, "-", Week_y),
                                   Comparison_y = paste0(Day_y,"-", Week_y, "_", Day_x, "-", Week_x),
                                   x_y = paste0(Condition, "_", variable),
                                   y_x = paste0(variable, "_", Condition),
                                   Comparison_xy = paste0(x_y, "_",y_x),
                                   Day_week_x = paste0(Day_x, "_", Week_x),
                                   Day_week_y = paste0(Day_y, "_", Week_y),
                                   Comparison = factor(Comparison, levels = day_levels)) %>% filter(Week_x == Week_y,
                                                                                                    Day_week_x != Day_week_y,
                                                                                                    Day_x != "D14",
                                                                                                    Day_y != "D10")


scaleFUN <- function(x) sprintf("%.2f", x)

p1 <- ggplot(plotDF2, aes(y=value))+
  geom_boxplot(aes(x=1), show.legend = F, outlier.shape = NA) +
  geom_jitter(aes(x=1, color=Comparison), show.legend = F, size = 2, width=0.05) +
  scale_color_viridis_d(name = "", option = "C") +
  scale_y_continuous(labels=scaleFUN) +
  facet_wrap(~"All Samples") +
  ylab("Jenson-Shannon Divergence\n") +
  theme_jsd1() %+replace% theme(legend.text = element_text(size=6),
                                             axis.text = element_text(size=6),
                                             strip.text = element_text(size = 6, margin = margin(0.1,0,0.1,0, "cm")),
                                             axis.title = element_text(size = 6))


p2 <- ggplot(plotDF2, aes(y=value))+
  geom_boxplot(aes(x=1), show.legend = F, outlier.shape = NA) +
  geom_jitter(aes(x=1, color=Comparison), show.legend = T, size = 2, width = 0.05) +
  scale_color_viridis_d(name = "", option = "C") +
  facet_wrap(~Week_x) +
  guides(color = guide_legend(ncol = 1)) +
  ylab("Jenson-Shannon Divergence") +
  theme_jsd2() %+replace% theme(legend.text = element_text(size=6),
                                             axis.text = element_text(size=6),
                                             strip.text = element_text(size = 6, margin = margin(0.1,0,0.1,0, "cm")),
                                             axis.title = element_text(size = 6))



# cowplot::plot_grid(plotlist = list(p1,p2,NULL), ncol = 3, rel_widths = c(2,3,1))



legend <- cowplot::get_legend(p2)

# p2 <- p2 + Seurat::NoLegend()

p2 <- p2 + theme(legend.position = "none")

# figs5a <- cowplot::plot_grid(plotlist = list(p1,p2,legend, NULL), ncol = 4, rel_widths = c(2,2,1,1))
# 
# figs5a

ab <- cowplot::plot_grid(plotlist = list(p1,p2), ncol = 2, rel_widths = c(3,3))
ab
figs5b <- cowplot::plot_grid(plotlist = list(ab,legend, NULL), ncol = 3, rel_widths = c(3.05,1,1))

figs5b

p1 <- p1 +
  ylim(ymin,ymax)

p2 <- p2 +
  ylim(ymin,ymax)

# cowplot::plot_grid(plotlist = list(p1,p2), ncol = 2, rel_widths = c(0.4,0.8))



```





```{r, fig.width=6}

plotDF2 <- plotDF2 %>% mutate(num_Day_x = as.numeric(str_split(Day_x,pattern = "D", simplify = T)[,2]),
                                        num_Day_y = as.numeric(str_split(Day_y,pattern = "D", simplify = T)[,2])) %>% filter(num_Day_x < num_Day_y)


p1 <- ggplot(plotDF2, aes(y=value))+
  geom_boxplot(aes(x=1), show.legend = F, outlier.shape = NA) +
  geom_jitter(aes(x=1, color=Comparison), show.legend = T, size = 2, width=0.05) +
  scale_color_viridis_d(name = "Comparison", option = "C") +
  #facet_wrap(~"All Samples") +
  ylab("Jenson-Shannon Divergence") +
  theme_jsd1()


p1

summaryDays <- plotDF2 %>% group_by(Comparison) %>% summarize(mean = mean(value), median = median(value), sd = sd(value))
summary <- plotDF2 %>% summarize(mean = mean(value), median = median(value), sd = sd(value))

```


```{r}
summary %>% kable() %>% kable_styling(full_width = F, position = "left")
summaryDays %>% kable() %>% kable_styling(full_width = F, position = "left")
```




## Distance Between Weeks (within days)


```{r fig.width=8, fig.height=5}

#### Fig S5A

# plotDF3 <- melt_sim_dis %>% mutate(Comparison = paste0(Day_x,"_", Week_x, "-", Day_y, "_", Week_y),
#                                    x_y = paste0(Condition, "-", variable),
#                                    y_x = paste0(variable, "-", Condition),
#                                    Comparison = factor(Comparison, levels = week_levels)) %>% filter(Day_x == Day_y)

# plotDF3 <- melt_sim_dis %>% mutate(Comparison = paste0(Day_x,"_", Week_x, "-", Day_y, "_", Week_y),
#                                    Comparison_y = paste0(Day_y,"_", Week_y, "-", Day_x, "_", Week_x),
#                                    x_y = paste0(Condition, "-", variable),
#                                    y_x = paste0(variable, "-", Condition),
#                                    Comparison_xy = paste0(x_y, "--",y_x),
#                                    Day_week_x = paste0(Day_x, "-", Week_x),
#                                    Day_week_y = paste0(Day_y, "-", Week_y),
#                                    Comparison = factor(Comparison, levels = day_levels)) %>% filter(Day_x == Day_y,
#                                                                                                     Week_x != "4wk",
#                                                                                                     Week_y != "2wk") %>% distinct(Comparison_xy, .keep_all = T)

plotDF3 <- melt_sim_dis %>% mutate(Comparison = paste0(Day_x,"-", Week_x, " vs ", Day_y, "-", Week_y),
                                   Comparison_y = paste0(Day_y,"-", Week_y, "_", Day_x, "-", Week_x),
                                   x_y = paste0(Condition, "_", variable),
                                   y_x = paste0(variable, "_", Condition),
                                   Comparison_xy = paste0(x_y, "_",y_x),
                                   Day_week_x = paste0(Day_x, "_", Week_x),
                                   Day_week_y = paste0(Day_y, "_", Week_y),
                                   Comparison = factor(Comparison, levels = day_levels)) %>% filter(Day_x == Day_y,
                                                                                       Week_x != "2wk",
                                                                                       Week_y != "4wk")


scaleFUN <- function(x) sprintf("%.2f", x)

p1 <- ggplot(plotDF3, aes(y=value))+
  geom_boxplot(aes(x=1), show.legend = F, outlier.shape = NA) +
  geom_jitter(aes(x=1, color=Comparison), show.legend = F, size = 2, width = 0.05) +
  scale_color_viridis_d(name = "", option = "C") +
  scale_y_continuous(labels=scaleFUN) +
  facet_wrap(~"All Samples") +
  ylab("Jenson-Shannon Divergence\n") +
  theme_jsd1() %+replace% theme(legend.text = element_text(size=6),
                                             axis.text = element_text(size=6),
                                             strip.text = element_text(size = 6, margin = margin(0.1,0,0.1,0, "cm")),
                                             axis.title = element_text(size = 6))


p2 <- ggplot(plotDF3, aes(y=value))+
  geom_boxplot(aes(x=1), show.legend = F, outlier.shape = NA) +
  geom_jitter(aes(x=1, color=Comparison), show.legend = T, size = 2, width = 0.05) +
  scale_color_viridis_d(name = "", option = "C") +
  facet_wrap(~Day_x) +
  guides(color = guide_legend(ncol = 1)) +
  ylab("Jenson-Shannon Divergence") +
  theme_jsd2() %+replace% theme(legend.text = element_text(size=6),
                                             axis.text = element_text(size=6),
                                             strip.text = element_text(size = 6, margin = margin(0.1,0,0.1,0, "cm")),
                                             axis.title = element_text(size = 6))


# cowplot::plot_grid(plotlist = list(p1,p2), ncol = 2, rel_widths = c(2,4))

legend <- cowplot::get_legend(p2)

# p2 <- p2 + Seurat::NoLegend()

p2 <- p2 + theme(legend.position = "none")

# figs5b <- cowplot::plot_grid(plotlist = list(p1,p2,legend), ncol = 3, rel_widths = c(2,3,1))
# 
# figs5b

ab <- cowplot::plot_grid(plotlist = list(p1,p2), ncol = 2, rel_widths = c(1,1.5))

figs5a <- cowplot::plot_grid(plotlist = list(ab,legend), ncol = 2, rel_widths = c(3,1))

figs5a

p1 <- p1 +
  ylim(ymin,ymax)

p2 <- p2 +
  ylim(ymin,ymax)

# cowplot::plot_grid(plotlist = list(p1,p2), ncol = 2, rel_widths = c(0.2,0.6))

summaryWeeks <- plotDF3 %>% group_by(Comparison) %>% summarize(mean = mean(value),median = median(value), sd = sd(value))
summaryWeeks2 <- plotDF3 %>% summarize(mean = mean(value),median = median(value), sd = sd(value))
```





```{r, fig.width=6}

p1 <- ggplot(plotDF3, aes(y=value))+
  geom_boxplot(aes(x=1), show.legend = F, outlier.shape = NA) +
  geom_jitter(aes(x=1, color=Comparison), show.legend = T, size = 2, width = 0.05) +
  scale_color_viridis_d(name = "Comparison", option = "C") +
  #facet_wrap(~"All Comparisons") +
  ylab("Jenson-Shannon Divergence") +
  theme_jsd1()

p1
```




## Fig S5

```{r fig.width=8, fig.height=10}

cowplot::plot_grid(figs5a, figs5b, ncol=1)

ggsave("journal_submission/figs5.png", device = "png", width = 8, height = 10, dpi = 600)

ggsave("journal_submission/figs5a.png", plot=figs5a, device = "png", width = 8, height = 5, dpi = 600)
ggsave("journal_submission/figs5b.png", plot=figs5b, device = "png", width = 8, height = 5, dpi = 600)

```



```{r}
summaryWeeks2 %>% kable() %>% kable_styling(full_width = F, position = "left")

summaryWeeks %>% kable() %>% kable_styling(full_width = F, position = "left")
```














## Cell type proportions

```{r}

in_vivo_cell_types %>% ungroup() %>%  mutate(Sample = paste(Condition, `Cell Type`, Sample, sep = "-"), Sample = gsub(" ", "_", Sample,fixed = T)) %>% 
  dplyr::select(Sample, Pct)

day_12 <- tibble(`Cell Type` = unique(c(in_vitro_cell_types$cell_type_character, in_vivo_cell_types$cell_type_character)))


in_vivo_cell_types$Sample <- as.factor(in_vivo_cell_types$Sample)


```









