---
title: "in vivo UMAPs"
output:
  html_document:
    code_folding: hide
    collapsed: no
    highlight: haddock
    keep_md: no
    theme: cerulean
    toc: yes
    toc_float:
      collapsed: yes
    toc_depth: 6
    number_sections: false
params:
    date: !r format(Sys.Date(), "%d %B, %Y")
    client: Goldfinch Bio

---

<!-- Indent the TOC -->
<script>
$(document).ready(function() {
  $items = $('div#TOC li');
  $items.each(function(idx) {
    num_ul = $(this).parentsUntil('#TOC').length;
    $(this).css({'text-indent': num_ul * 10, 'padding-left': 0});
  });

});
</script>

<table width=100%>
<tr>
<td>
<h3>Prepared for `r params$client` by:</h3>
Diamond Age Data Science
</td>
<td width=150></td>
<td width=150 align=right>`r params$date`</td>
</tr>
</table>




```{r setup, include=FALSE}
DOCNAME = "in_vivo_UMAPs"
knitr::opts_chunk$set(autodep        = TRUE,
                      cache          = FALSE,
                      cache.path     = paste0("cache/", DOCNAME, "/"),
                      fig.path       = paste0("cache/", DOCNAME, "/"),
                      cache.comments = FALSE,
                      cache.lazy     = FALSE,
                      echo           = TRUE,
                      error          = FALSE,
                      fig.align      = "center",
                      fig.width      = 10,
                      fig.height     = 8,
                      message        = FALSE,
                      warning        = FALSE)

#Libraries
library(Seurat)
library(parallel)
library(future)
library(here)
library(GSVA)
library(clustree)
library(viridis)
library(knitr)
library(DT)
library(kableExtra)
library(tidyverse)

source(here("R/utilities.R"))
rename = dplyr::rename


#Options
nCores <- 18
options(future.globals.maxSize = 16 * 1024^3)
save = TRUE



# Directories
baseDir = here()
saveDir = file.path(baseDir,"output", DOCNAME)
if(!dir.exists(saveDir)){
  dir.create(saveDir)
}



```





```{r}
conditions <- c("Day_10_2wk_transplant", "Day_10_4wk_transplant", "Day_12_2wk_transplant", "Day_12_4wk_transplant", "Day_14_2wk_transplant", "Day_14_4wk_transplant")
names(conditions) <- c("D10-2wk", "D10-4wk", "D12-2wk", "D12-4wk", "D14-2wk", "D14-4wk")

collapse <- c("S-Shaped Body", "Comma-Shaped Body", "Renal Vessicle", "Cap Mesenchyme")


seur_list <- lapply(conditions, function(c){
    seur <- read_rds(file.path(baseDir,"output",c,paste0(c, ".RDS")))
    seur$cell_type_simple_2 <- factor(seur$cell_type_simple_2, levels = cell_levels_collapse)
    return(seur)
    # temp_meta <- tibble(cell_name = rownames(seur@meta.data),
    #                 cell_type_simple = seur$cell_type_simple_2)
    # 
    # temp_meta <- temp_meta %>% mutate(cell_type_simple = ifelse(cell_type_manual %in% collapse, "Nephron Progenitor", cell_type_manual),
    #                                   cell_type_simple = factor(cell_type_simple, levels = cell_levels_collapse)) %>% 
    # select(cell_name, cell_type_simple) %>% column_to_rownames("cell_name")
    # 
    # seur <- AddMetaData(seur, temp_meta)
})


```

# ***in vivo*** Plots


## combined UMAP panels

```{r}

plots <- lapply(1:length(conditions), function(x){
                DimPlot(seur_list[[x]], reduction = 'umap',
                        group.by = 'cell_type_simple_2', pt.size = 0.1) + 
                        scale_color_manual(values = colors, drop = FALSE) +
                        ggtitle(names(conditions)[x]) +
                                      theme(legend.text = element_text(size=8),
                                            legend.spacing = unit(10, "points"),
                                            legend.key.height = unit(8, "points"),
                                            axis.title = element_text(size=14),
                                            plot.title = element_text(hjust = 0.5, size = 14))
})

plots_no_legend <- lapply(plots, function(x){x + NoLegend()})


```


```{r, fig.height=4, fig.width=3}
legend <- cowplot::get_legend(plots[[1]]+
                                theme(legend.text = element_text(size=14),
                                      legend.spacing = unit(16, "points"),
                                      legend.key.height = unit(14, "points")))

```


```{r, fig.height=16, fig.width=12}
p1 <- cowplot::plot_grid(plotlist = plots_no_legend, labels = "AUTO", ncol = 2)

cowplot::plot_grid(p1, legend, ncol = 2, rel_widths = c(3.5,1))


```



## Individual UMAPs {.tabset}



```{r cluster-umap, results = "hide"}

src_list <- lapply(1:length(conditions), function(c) {
    src <- c("### `r names(conditions)[{{c}}]` {.unnumbered}",
             "```{r labeled-umap-{{c}}}",
             "DimPlot(seur_list[[{{c}}]], reduction = 'umap',
                      group.by = 'cell_type_simple_2', pt.size = 0.3) +
                      scale_color_manual(values = colors, drop = FALSE) +
                      ggtitle(names(conditions)[{{c}}]) +
                      theme(legend.text = element_text(size=14),
                            axis.title = element_text(size=14),
                            plot.title = element_text(hjust = 0.5))",  
             "```",
             "")
    knit_expand(text = src)
})

out <- knit_child(text = unlist(src_list), options = list(cache = FALSE))
```

`r out`



## TRPC expression  {.tabset}



```{r feature-umap, results = "hide"}

src_list <- lapply(1:length(conditions), function(c) {
    src <- c("### `r names(conditions)[{{c}}]` {.unnumbered}",
             "```{r feature-{{c}}}",
             "FeaturePlot(seur_list[[{{c}}]], reduction = 'umap',
                      features = c('TRPC4','TRPC5','TRPC6'),min.cutoff='q9') +
                      ggtitle(names(conditions)[{{c}}])",  
             "```",
             "")
    knit_expand(text = src)
})

out <- knit_child(text = unlist(src_list), options = list(cache = FALSE))
```

`r out`


## RET expression  {.tabset}



```{r ret-feature-umap, results = "hide"}

src_list <- lapply(1:length(conditions), function(c) {
    src <- c("### `r names(conditions)[{{c}}]` {.unnumbered}",
             "```{r ret-feature-{{c}}}",
             "FeaturePlot(seur_list[[{{c}}]], reduction = 'umap',
                      features = c('AQP2', 'CLDN4','GATA3', 'ELF5'),min.cutoff='q9') +
                      ggtitle(names(conditions)[{{c}}])",  
             "```",
             "")
    knit_expand(text = src)
})

out <- knit_child(text = unlist(src_list), options = list(cache = FALSE))
```

`r out`


```{r}

```





